---
title: "Case Study 1 - Cyclistic Marketing"
author: "Salimah Ismail"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Case Study 1: How Does a Bike-Share Navigate Speedy Success?

Cyclistic is a fictional bike-share company in Chicago who has several pricing plans: single ride passes, full-day passes and annual memberships. Customers who purchase single rides or full-day passes are considered casual members, whereas those who have annual memberships are considered Cyclistic members. 

## The Task
Cyclistic believes that there is a growth opportunity in converting casual riders into full Cyclistic members. This analysis aims to determine **the main differences between how casual members use Cyclistic versus annual members** so that the Cyclistic marketing department can understand these differences and market annual memberships to these casual customers effectively. 

### The Data
Cyclistic data sets for analysis are available [here](https://divvy-tripdata.s3.amazonaws.com/index.html), mainly organized as one file per month. I chose to examine one year of data, from July 2022 - 2023 to get the most recent insights on riders.  


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
